### 2018-06-20 paper/4-build.tex
Wibbles to start of 5

With the information in this section we can build a table comparing the
dependency order strategy with the out-of-date keys strategy, providing 9~possible
build systems, 5~of which are actually inhabited by existing build systems
(we discuss \Ninja \cite{ninja} in \S\ref{sec-related-build}). Of the
remaining 4~spots, we believe neither Traced or Cloud \Excel make sense~--~the
\Excel approach of reordering combined with static approximations reduces
the memory usage significantly. However, as soon as you are paying the cost of
storing traces, that benefit is gone. The advantage of Approximate \Shake over
\Shake would be that it could avoid storing traces and having a separate
information database, but that advantage is minor compared to the technical
restrictions and approximations it would provide, so we consider it unlikely
to be built. The Cloud \Shake system is an interesting and important point in
the design space, which we explore further
in~\S\ref{sec-implementation-cloud}.

>>>>

With the information in this section we can build a table comparing the
dependency order strategy with the out-of-date keys strategy, providing 9~possible
build systems \simon{Or perhaps 12 now?}, 5~of which are actually inhabited by existing build systems
(we discuss \Ninja \cite{ninja} in \S\ref{sec-related-build}). Of the
remaining 4~spots, we believe neither Traced or Cloud \Excel make sense~--~the
\Excel approach of reordering combined with static approximations reduces
the memory usage significantly. However, as soon as you are paying the cost of
storing traces, that benefit is gone. The advantage of Approximate \Shake over
\Shake would be that it could avoid storing traces and having a separate
information database, but that advantage is minor compared to the technical
restrictions and approximations it would provide, so we consider it unlikely
to be built. The Cloud \Shake system is an interesting and important point in
the design space, which we explore further
in~\S\ref{sec-implementation-cloud}.

### 2018-06-20 paper/5-implementations.tex
Wibbles to start of 5

In the previous sections we discussed the types of build systems, and how they
can be broken down into two main components: a scheduler and a rebuilder.
In this section we make this abstract distinction concrete, by actually
implementing a number of build systems as a composition of a schedulear
and a rebuilder.  Our design is expressed by these types:
\begin{minted}[xleftmargin=10pt]{haskell}
type Scheduler c i k v = Rebuilder c i k v -> Build c k v
type Rebuilder c i k v = k -> v -> Task c k v -> Task (MonadState i) k v
\end{minted}
That is:
\begin{itemize}
\item A \hs{Scheduler} is just a function that takes a \hs{Rebuilder} and uses
  it to construct \hs{Build} system, by choosing which keys to rebuild in which order.
\item A \hs{Rebuilder} is a function that is given a key, the key's ``old'' value,
  and a \hs{Task} that can compute the value of the key if necessary. It uses
  the persistent build state \hs{i} (carried by the state monad) to decide
  whether to re-build the value.  If doing so is unnecesary, it returns the
  old value; otherwise it runs the supplied \hs{Task} to rebuild it.  In both
  cases it updates the persistent build information \hs{i} to reflect what happened.
\end{itemize}

@@

\Make's scheduler processes keys in a linear \hs{order} based on a
\hs{topological} sort of the statically known dependency graph
(see~\S\ref{sec-parallelism} for parallel \Make). Our definition in
Fig.~\ref{fig-make-implementation} is polymorphic with respect to the type of
build information \hs{i} and is therefore compatible with any applicative
\hs{rebuilder}. The scheduler calls the supplied \hs{rebuilder} on every
\hs{key} in the \hs{order}, and runs the obtained \hs{newTask} to compute the
\hs{newValue} and updated build information \hs{newInfo}, which are subsequently
written to the store. The computation runs in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v)} monad, similar to the \hs{busy}
build system (\S\ref{sec-general-build}), however since \hs{newTask} has access
only to the \hs{i} part of the \hs{Store}~\hs{i}~\hs{k}~\hs{v}, we need to use
\hs{runState} to execute it. The \hs{newTask} finds values of the \hs{key}'s
dependencies via the \hs{fetch} callback, which is defined to directly read the
\hs{store}.

>>>>

In the previous sections we discussed the types of build systems, and how they
can be broken down into two main components: a scheduler and a rebuilder.
In this section we make this abstract distinction concrete, by
implementing a number of build systems as a composition of a scheduler
and a rebuilder.  Our design is expressed by these types:
\begin{minted}[xleftmargin=10pt]{haskell}
type Scheduler c i k v = Rebuilder c i k v -> Build c k v
type Rebuilder c i k v = k -> v -> Task c k v -> Task (MonadState i) k v
\end{minted}
That is:
\begin{itemize}
\item A \hs{Scheduler} is just a function that takes a \hs{Rebuilder} and uses
  it to construct \hs{Build} system, by choosing which keys to rebuild in which order.
\item A \hs{Rebuilder} is a function that is given a key, the key's ``old'' value,
  and a \hs{Task} that can compute the value of the key if necessary. It uses
  the persistent build state \hs{i} (carried by the state monad) to decide
  whether to re-build the value.  If doing so is unnecesary, it returns the
  old value; otherwise it runs the supplied \hs{Task} to rebuild it.  In both
  cases it updates the persistent build information \hs{i} to reflect what happened.
\end{itemize}
We write a scheduler for each column of Table~\ref{tab-build-systems}, and a
rebuilder for each row; then applying the former to the latter, in any combination,
gives a correct build system.

@@

\Make's scheduler, \hs{topological}, processes keys in a linear \hs{order} based on a
topological sort of the statically known dependency graph
(see~\S\ref{sec-parallelism} for parallel \Make). Our definition in
Fig.~\ref{fig-make-implementation} is polymorphic with respect to the type of
build information \hs{i} and is therefore compatible with any applicative
\hs{rebuilder}. The scheduler calls the supplied \hs{rebuilder} on every
\hs{key} in the \hs{order}, and runs the obtained \hs{newTask} to compute the
\hs{newValue} and updated build information \hs{newInfo}, which are subsequently
written to the store. The computation runs in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v)} monad, similar to the \hs{busy}
build system (\S\ref{sec-general-build}), however since \hs{newTask} has access
only to the \hs{i} part of the \hs{Store}~\hs{i}~\hs{k}~\hs{v}, we need to use
\hs{runState} to execute it. The \hs{newTask} finds values of the \hs{key}'s
dependencies via the \hs{fetch} callback, which is defined to directly read the
\hs{store}.
